
---
title: "Mixture Model Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document demonstrates how to analyze a dataset using mixture models. The dataset contains variables like performance, stress levels, age, and gender. We will identify latent subgroups (classes) in the data using mixture modeling techniques.

---

## Step 1: Load the Data

```{r load-data}
# Load necessary libraries
library(tidyverse)

# Load the dataset (assuming it's in the same folder as this Rmd file)
data <- read.csv("mixture_data.csv")

# Quick look at the data
head(data)
summary(data)
```

---

## Step 2: Visualize the Data

```{r visualize-data}
# Distribution of performance
ggplot(data, aes(x = Performance)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "black") +
  labs(title = "Performance Distribution", x = "Performance", y = "Count")

# Distribution of stress
ggplot(data, aes(x = Stress)) +
  geom_histogram(binwidth = 1, fill = "orange", color = "black") +
  labs(title = "Stress Distribution", x = "Stress Level", y = "Count")

# Scatterplot: Performance vs Stress
ggplot(data, aes(x = Stress, y = Performance, color = Gender)) +
  geom_point(size = 3) +
  labs(title = "Performance vs Stress", x = "Stress Level", y = "Performance")
```

---

## Step 3: Fit a Mixture Model

```{r fit-mixture-model, message=FALSE, warning=FALSE}
# Load the mixture modeling library
library(mclust)

# Fit a mixture model
mixture_model <- Mclust(data[, c("Performance", "Stress")])

# Summary of the mixture model
summary(mixture_model)

# Plot the mixture model results
plot(mixture_model, what = "classification")
```

---

## Step 4: Interpret Results

```{r interpret-results}
# Display the classification
data$Class <- mixture_model$classification
table(data$Class)

# Visualize the classes
ggplot(data, aes(x = Stress, y = Performance, color = factor(Class))) +
  geom_point(size = 3) +
  labs(title = "Mixture Model Classes", x = "Stress Level", y = "Performance", color = "Class")
```

---

## Conclusion

This document provides a simple framework for mixture model analysis. The dataset has been split into latent classes based on performance and stress. You can extend this analysis by incorporating more variables or testing alternative model structures.
